Measuring causal influence with\\ back-to-back regression: the linear case

### 2019-11-12 paper/introduction.tex
correct alex comments

The second challenge to measuring causal influence is that observations can be large and complex. The relationship between causes and effects is thus often considered in a backward manner, by training models to maximally predict causes from multidimensional observations. For example, brain activity is often recorded with hundreds or thousands of sensors simultaneously. As multiple sensors may be affected by common noise sources, it is common to use `decoding' techniques, by, for example, fitting a support vector machine across multiple sensors to predict the category of a stimulus \citep{cichy2014resolving,  kriegeskorte2008representational, norman2006beyond}.

Both forward and backward modeling have competing benefits and drawbacks. Specifically, forward modeling disentangles the independent contribution of collinear factors, but does not combine multidimensional observations. By contrast, backward modeling combines multiple observations, but does not disentangle collinear factors \cite{weichwald2015causal, hebart2018deconstructing, king2018encoding}. To combine the benefits of forward and backward modeling, several authors have proposed to use cross-decomposition techniques such as Partial Least Squares (PLS) and Canonical Correlation Analysis (CCA) \citep{de2019multiway}.
CCA and PLS aim to find, from two sets of data $X$ and $Y$, the components $H$ and $G$ were $XH$ and $YG$ are maximially correlated or maximally covarying respectively. Because CCA and PLS are are based on a generalized eigen decomposition, their resulting coefficients are mixing the features of $X$ and $Y$ in a way that makes them notoriously difficult to interpret \citep{lebart1995statistique}.

>>>>

The second challenge to measuring causal influence is that observations can be large and complex. The relationship between causes and effects is thus often considered in a backward manner, by training models to maximally predict causes from multidimensional observations. For example, brain activity is often recorded with hundreds or thousands of sensors simultaneously. As multiple sensors may be affected by common noise sources, it is common to use backward regression, by, for example, fitting a support vector machine across multiple sensors to decode the category of a stimulus \citep{cichy2014resolving,  kriegeskorte2008representational, norman2006beyond}.

Both forward and backward modeling have competing benefits and drawbacks. Specifically, forward modeling disentangles the independent contribution of collinear factors, but does not combine multidimensional observations. By contrast, backward modeling combines multiple observations, but does not disentangle collinear factors \citep{weichwald2015causal, hebart2018deconstructing, king2018encoding}. To combine the benefits of forward and backward modeling, several authors have proposed to use cross-decomposition techniques such as Partial Least Squares (PLS) and Canonical Correlation Analysis (CCA) \citep{de2019multiway}.
CCA and PLS aim to find, from two sets of data $X$ and $Y$, the components $H$ and $G$ were $XH$ and $YG$ are maximially correlated or maximally covarying respectively. Because CCA and PLS are based on a generalized eigen decomposition, their resulting coefficients are mixing the features of $X$ and $Y$ in a way that makes them notoriously difficult to interpret \citep{lebart1995statistique}.

### 2019-11-12 paper/paper.tex
correct alex comments

\begin{figure}[t!]
    \centering
    \begin{tikzpicture}
    \newcommand\posY{0}
    \newcommand\posX{3}
    \newcommand\posE{5}
    \newcommand\posN{7}
    \newcommand\posF{9.5}

@@

    \node[] (plus) at (6, 0){$+$};
    \node[thick, draw=black, minimum height=3cm, minimum width=1cm, fill=blue!10] (N) at (\posN, 0){};
    \node[thick, draw=black, minimum height=2cm, minimum width=1cm, fill=red!10] (F) at (\posF, 0){};
    \draw [thick] (2.25, -1.5) to [round left paren ] (2.25, 1.5);
    \draw [thick] (7.75, -1.5) to [round right paren ] (7.75, 1.5);

@@

    \node[] (sim1) at (0,-3.25) {\scalebox{0.85}{$X \sim P(X)$}};
    \node[] (sim2) at (0,-3.65) {\scalebox{0.85}{$N \sim P(N)$}};
    \node[] (reg1) at (5.5,-3.25) {$\hat{E} = \text{Diag}(\underbrace{(X_2^\top X_2 + \Lambda_X)^{-1} X_2^\top Y_2\overbrace{(Y_1^\top Y_1 + \Lambda_Y)^{-1} Y_1^\top X_1}^{\text{1) } \hat{X} : \text{ regression from } Y \text{ to } X}}_{\text{2) regression from } X \text{ to } \hat{X}})$};
    \end{tikzpicture}
    \caption{Back-to-back regression identifies the subset of factors $E_{ii} = 1$ in $X$ that influence some observations $Y$ by 1) regressing from $Y$ to $X$ to obtain $\hat{X}$, and ii) returning the diagonal of the regression coefficients from $X$ to $\hat{X}$.}
    \label{fig:b2b}
\end{figure}

We consider the measurement of multivariate signal $Y$, generated from a set of putative causes $X$, via some unknown linear apparatus $F$.
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $E$, we denote by $XE$ the causal factors of $Y$.
%
In summary, the problem can be formalized as:
%
\begin{equation}
    Y = (XE + N)F,\label{eq:model}
\end{equation}
%
where $N$ is a centered, unobserved noise variable.
%
While the triplet of variables $X$ and $N$ are independent, we allow each of them to have any form of covariance.
% what is a general covariance matrix?
%
In practice, we observe $n$ samples $(X, Y)$ from the model.
%
This problem space, along with the sizes of all variables involved, is illustrated in Figure~\ref{fig:b2b}.
%
Given the model in Equation~\eqref{eq:model}, \textbf{the goal} of Back-to-Back Regression (B2B) is to estimate the matrix of causal influences $E$.

@@

Back-to-Back Regression (B2B) consists of two steps.
%
First, we estimate the linear regression coefficients $\hat G$ from $Y$ to $X$, and construct the predictions $\hat X = Y \hat G$.
%
This backward regression recovers the correlations between $Y$ and each factor of $X$.
%
Second, we estimate the linear regression coefficients $\hat H$ from $X$ to $\hat X$.
%
The diagonal of the regression coefficients $\hat H$, denoted by $\hat{E} = \text{Diag}(\hat{H})$, is the desired estimate of the causal influence matrix $E$.

@@

Performing two regressions over the same data sample can result in overfitting, as spurious correlations in the data absorbed by the first regression will be leveraged by the second one.
%
To avoid this issue, we split our sample $(X, Y)$ into two splits $(X_1, Y_1)$ and $(X_2, Y_2)$.
%
Then, the first regression is performed using $(X_1, Y_1)$, and the second regression is performed using $(X_2, Y_2)$.
%
To compensate for the reduction in sample size caused by the split, B2B is repeated over many random splits, and the final estimate $\hat E$ of the causal influence matrix is the average over the estimates associated to each split \citep{efron1992bootstrap}.
%
To accelerate this ensembling procedure, we implemented an efficient leave-one-out cross-validation scheme as detailed in \citep{rifkin2007notes} % http://cbcl.mit.edu/publications/ps/MIT-CSAIL-TR-2007-025.pdf
, as follows:
%

\begin{equation}
\hat{Y} = (\Sigma_X G Y - diag(\Sigma_X G) Y) \;/\; diag(I - \Sigma_X G) \qquad \text{(element-wise division)}
\end{equation}

where $\Sigma_X$ is the $X$ kernel matrix and where $G$ is computed with an eigen decomposition of $X$:
\begin{equation}
  \begin{aligned}
  \Sigma_X = Q V Q^T \\
  G = Q (V + \lambda*I)^{-1} Q^T\\
  \end{aligned}
\end{equation}
where $Q$, $V$ and $\lambda$ are the eigen vectors, eigen values and regularization, respectively.

@@

% JR: shall we update algorithm with LOO trick?
\begin{algorithm}[H]
    %\SetAlgoLined
    \KwIn{input data $X \in \mathbb{R}^{n \times d_x}$, output data $Y \in \mathbb{R}^{n\times d_y}$, number of repetitions $m \in \mathbb{N}$.}
    \KwOut{estimate of causal influences $\hat{E} \in \mathbb{D}^{d_x \times d_x}$.}
    $\hat{E} \leftarrow 0$\;
    \For{$i = 1, \ldots, m$}{
        $(X, Y) \leftarrow \text{ShuffleRows}((X, Y))$\;
        $(X_1, Y_1), (X_2, Y_2) \leftarrow \text{SplitRowsInHalf}((X, Y))$\;
        $\hat{G} = \text{LinearRegression}(Y_1, X_1)$ \Comment*[r]{$\hat G = (Y_1^\top Y_1 + \Lambda_Y)^{-1} Y_1^\top X_1$}
        $\hat{H} = \text{LinearRegression}(X_2, Y_2 \hat{G})$ \Comment*[r]{$\hat H=(X_2^\top X_2 + \Lambda_X)^{-1} X_2^\top Y_2 \hat G$}
        $\hat{E} \leftarrow \hat{E} + \text{Diag}(\hat{H})$\;
    }
    $\hat{E} \leftarrow \hat{E} / m$\;
    $\hat{W} \leftarrow \text{LinearRegression}(X \hat{E}, Y)$\;
    \Return{$\hat{E}$, $\hat{W}$}
    \caption{Back-to-back regression.}
    \label{algorithm:b2br}
\end{algorithm}

@@

     Consider the B2B model from Equation $Y = (XE + N)F$, $N$ centred and full rank noise.
     %
     If $F$ and $X$ are full-rank on $Img(E)$, then, the solution of B2B, $\hat H$, will minimize
     %
     $\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$and satisfy $E\hat H = \hat H$
\end{theorem}
%
\begin{proof}
  See Appendix \ref{appendix:theorem_proof}.
\end{proof}

@@

Assuming, without loss of generality, that the active features in $E$ are the $k$ first, and rewriting $X=(X_1,X_2)$ and $N=(N_1,N_2)$ ($X_1$ and $N_1$ containing the $k$ first features), we have ($\Sigma_{A B}$ denoting the covariance of $A$ and $B$)

% $$
\begin{equation}
  X^\top X = \left(\begin{array}{lccl}\Sigma_{X_1 X_1} & \Sigma_{X_1 X_2} \\ \Sigma_{X_1 X_2} & \Sigma_{X_2 X_2}\end{array}\right),\qquad\qquad N^\top N = \left(\begin{array}{lccl}\Sigma_{N_1 N_1} & \Sigma_{N_1 N_2} \\ \Sigma_{N_1 N_2} & \Sigma_{N_2 N_2}\end{array}\right),
\end{equation}
% $$
% and
\begin{equation}
  \hat H = \left(\begin{array}{cc}(\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_1} & (\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_2} \\0 & 0\end{array}\right),
\end{equation}
\begin{equation}
  Diag_k (\hat H) = Diag((\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_1}) = Diag((I+\Sigma_{X_1 X_1}^{-1}\Sigma_{N_1 N_1})^{-1}).
\end{equation}
%
% $$
% $$
% $$
% $$
%
In the absence of noise, we have $\Sigma_{N_1 N_1}=0$, and so $Diag_k(\hat H)=I$, and $$Diag(\hat H) = Diag(E)$$ 
Otherwise the $k$ first diagonal elements of $\hat H$ are all positive (since we assume that $F$ has full rank over $Img(E)$, they cannot be zero), lower than $1$, and bounded by 
$$\frac{\sigma_{X_k}}{\sigma_{X_k} +\sigma_{N_1}} \text{ and }  \frac{\sigma_{X_1}}{\sigma_{X_1} +\sigma_{N_k}}$$where  $\sigma_{X_1}$, $\sigma_{X_k}$, $\sigma_{N_1}$ and $\sigma_{N_k}$ denote the largest and smallest eigenvalues of $\Sigma_{X_1 X_1}$ and $\Sigma_{N_1 N_1}$.

In practical cases, non linearities and sampling error will add noise to the diagonal elements of $\hat H$. To test diagonal elements for equality with zero, we note that the average value of the non-zero coefficients of $Diag(\hat H)$ is the trace of $\hat H$ divided by $k$. Since $X$ and $N$ are not correlated, $\Sigma_{X_1 X_1}^{-1}\Sigma_{N_1 N_1}$ should have a low condition number, which means $\frac{Var(X)}{Var(X)+Var(N)}$ should provide a good estimate of $\frac{1}{k}Tr(\hat H)$. This can be used to test whether a feature of $X$ has causal influence. An alternative is the Sonquist-Morgan criterion presented in the appendix.

@@

Finally, B2B is also related to, but different from causal discovery algorithms used outside the neuroimaging field. These methods are often nonlinear \citep{peters2017elements} and base their causal estimates on temporal delays \citep{granger1969investigating, janzing2013quantifying}. For example, B2B relates to tthe half-sibling regression \citep{scholkopf2016modeling}, in the sense that we observe a large number of variables corrupted by the same source of noise; we leave the extension of B2B to nonlinear time-series for future work.

>>>>

\begin{figure}[t!]
    \centering
    \begin{tikzpicture}
    \newcommand\posY{0}
    \newcommand\posX{3}
    \newcommand\posE{5}
    \newcommand\posN{7}
    \newcommand\posF{10}

@@

    \node[] (plus) at (6, 0){$+$};
    \node[thick, draw=black, minimum height=3cm, minimum width=1cm, fill=blue!10] (N) at (\posN, 0){};
    \node[thick, draw=black, minimum height=1cm, minimum width=2cm, fill=red!10] (F) at (\posF, 0){};
    \draw [thick] (2.25, -1.5) to [round left paren ] (2.25, 1.5);
    \draw [thick] (7.75, -1.5) to [round right paren ] (7.75, 1.5);

@@

    \node[] (sim1) at (0,-3.25) {\scalebox{0.85}{$X \sim P(X)$}};
    \node[] (sim2) at (0,-3.65) {\scalebox{0.85}{$N \sim P(N)$}};
    \node[] (reg1) at (5.5,-3.25) {$\hat{E} = \text{diag}(\underbrace{(X_2^\top X_2 + \Lambda_X)^{-1} X_2^\top Y_2\overbrace{(Y_1^\top Y_1 + \Lambda_Y)^{-1} Y_1^\top X_1}^{\text{1) } \hat{X} : \text{ regression from } Y \text{ to } X}}_{\text{2) regression from } X \text{ to } \hat{X}})$};
    \end{tikzpicture}
    \caption{Back-to-back regression identifies the subset of factors $E_{ii} = 1$ in $X$ that influence some observations $Y$ by 1) regressing from $Y$ to $X$ to obtain $\hat{X}$, and 2) returning the diagonal of the regression coefficients from $X$ to $\hat{X}$.}
    \label{fig:b2b}
\end{figure}

We consider the measurement of multivariate signal $Y \in \mathbb{R}^{n \times d_y}$, generated from a set of putative causes $X \in \mathbb{R}^{n \times d_x}$, via some unknown linear apparatus $F \in \mathbb{R}^{d_x \times d_y}$.
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $E \in \mathbb{D}^{d_x \times d_x}$, we denote by $XE$ the causal factors of $Y$.
%
In summary, the problem can be formalized as:
%
\begin{equation}
    y_i = (x_i E + n_i) F
    \label{eq:model}
\end{equation}
%
where $i$ is a given sample, and $n$ is a centered, unobserved noise variable.
%
While the triplet of variables $X$ and $N$ are independent, we allow each of them to have any form of covariance.
% what is a general covariance matrix?
%
In practice, we observe $n$ samples $(X, Y)$ from the model.
%
This problem space, along with the sizes of all variables involved, is illustrated in Figure~\ref{fig:b2b}.
%
Given the model in Equation~\eqref{eq:model}, \textbf{the goal} of Back-to-Back Regression (B2B) is to estimate the matrix of causal influences $E$.

@@

Back-to-Back Regression (B2B) consists of two steps.
%
First, we estimate the linear regression coefficients $\hat G$ from $Y$ to $X$, and construct the predictions $\hat X = Y \hat G$.
%
This backward regression recovers the correlations between $Y$ and each factor of $X$.
%
Second, we estimate the linear regression coefficients $\hat H$ from $X$ to $\hat X$.
%
The diagonal of the regression coefficients $\hat H$, denoted by $\hat{E} = \text{diag}(\hat{H})$, is the desired estimate of the causal influence matrix $E$.

@@

Performing two regressions over the same data sample can result in overfitting, as spurious correlations in the data absorbed by the first regression will be leveraged by the second one.
%
To avoid this issue, we split our sample $(X, Y)$ into two splits $(X_1, Y_1)$ and $(X_2, Y_2)$.
%
Then, the first regression is performed using $(X_1, Y_1)$, and the second regression is performed using $(X_2, Y_2)$.
%
To compensate for the reduction in sample size caused by the split, B2B is repeated over many random splits, and the final estimate $\hat E$ of the causal influence matrix is the average over the estimates associated to each split \citep{breiman1996bagging}.
%
To accelerate this ensembling procedure, we implemented an efficient leave-one-out cross-validation scheme as detailed in \citep{rifkin2007notes} % http://cbcl.mit.edu/publications/ps/MIT-CSAIL-TR-2007-025.pdf
, as follows:
%

\begin{equation}
\hat{Y} = (\Sigma_X G Y - \text{diag}(\Sigma_X G) Y) \;/\; \text{diag}(I - \Sigma_X G) \qquad \text{(element-wise division)}
\end{equation}

where $\Sigma_X$ is the $X$ kernel matrix and where $G$ is computed with an eigen decomposition of $X$:
\begin{equation}
  \begin{aligned}
  \Sigma_X = Q V Q^T \\
  G = Q (V + \lambda I)^{-1} Q^T\\
  \end{aligned}
\end{equation}
where $Q$, $V$ and $\lambda$ are the eigen vectors, eigen values and regularization, respectively.

@@

% JR: shall we update algorithm with LOO trick?
\begin{algorithm}[H]
    %\SetAlgoLined
    \KwIn{input data $X \in \mathbb{R}^{n \times d_x}$, output data $Y \in \mathbb{R}^{n\times d_y}$, number of repetitions $m \in \mathbb{N}$.}
    \KwOut{estimate of causal influences $\hat{E} \in \mathbb{D}^{d_x \times d_x}$.}
    $\hat{E} \leftarrow 0$\;
    \For{$i = 1, \ldots, m$}{
        $(X, Y) \leftarrow \text{ShuffleRows}((X, Y))$\;
        $(X_1, Y_1), (X_2, Y_2) \leftarrow \text{SplitRowsInHalf}((X, Y))$\;
        $\hat{G} = \text{LinearRegression}(Y_1, X_1)$ \Comment*[r]{$\hat G = (Y_1^\top Y_1 + \Lambda_Y)^{-1} Y_1^\top X_1$}
        $\hat{H} = \text{LinearRegression}(X_2, Y_2 \hat{G})$ \Comment*[r]{$\hat H=(X_2^\top X_2 + \Lambda_X)^{-1} X_2^\top Y_2 \hat G$}
        $\hat{E} \leftarrow \hat{E} + \text{diag}(\hat{H})$\;
    }
    $\hat{E} \leftarrow \hat{E} / m$\;
    $\hat{W} \leftarrow \text{LinearRegression}(X \hat{E}, Y)$\;
    \Return{$\hat{E}$, $\hat{W}$}
    \caption{Back-to-back regression.}
    \label{algorithm:b2br}
\end{algorithm}

@@

     Consider the B2B model from Equation $Y = (XE + N)F$, $N$ centred and full rank noise.
     %
     Let $Img(M)$ refers to the image of the matrix $M$. If $F$ and $X$ are full-rank on the $Img(E)$, then, the solution of B2B, $\hat H$, will minimize
     %
     $\min_H  \left \| X - XH\right\| ^2  + \left \| NH\right \| ^2$and satisfy $E\hat H = \hat H$
\end{theorem}
%
\begin{proof}
  See Appendix \ref{appendix:theorem_proof}.
\end{proof}

@@

Assuming, without loss of generality, that the active features in $E$ are the $k \in \mathbb{Z}: k \in [0, d_x]$ first features, and rewriting $X=(X_1,X_2)$ and $N=(N_1,N_2)$ ($X_1$ and $N_1$ containing the $k$ first features), we have ($\Sigma_{A B}$ denoting the covariance of $A$ and $B$)

% $$
\begin{equation}
  X^\top X = \left(\begin{array}{lccl}\Sigma_{X_1 X_1} & \Sigma_{X_1 X_2} \\ \Sigma_{X_1 X_2} & \Sigma_{X_2 X_2}\end{array}\right),\qquad\qquad N^\top N = \left(\begin{array}{lccl}\Sigma_{N_1 N_1} & \Sigma_{N_1 N_2} \\ \Sigma_{N_1 N_2} & \Sigma_{N_2 N_2}\end{array}\right),
\end{equation}
% $$
% and
\begin{equation}
  \hat H = \left(\begin{array}{cc}(\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_1} & (\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_2} \\0 & 0\end{array}\right),
\end{equation}
\begin{equation}
  \text{diag}_k (\hat H) = \text{diag}((\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_1}) = \text{diag}((I+\Sigma_{X_1 X_1}^{-1}\Sigma_{N_1 N_1})^{-1}).
\end{equation}
%
% $$
% $$
% $$
% $$
%
In the absence of noise, we have $\Sigma_{N_1 N_1}=0$, and so $\text{diag}_k(\hat H)=I$, and $$\text{diag}(\hat H) = \text{diag}(E)$$
Otherwise the $k$ first diagonal elements of $\hat H$ are all positive (since we assume that $F$ has full rank over $Img(E)$, they cannot be zero), lower than $1$, and bounded by
$$\frac{\sigma_{X_k}}{\sigma_{X_k} +\sigma_{N_1}} \text{ and }  \frac{\sigma_{X_1}}{\sigma_{X_1} +\sigma_{N_k}}$$where  $\sigma_{X_1}$, $\sigma_{X_k}$, $\sigma_{N_1}$ and $\sigma_{N_k}$ denote the largest and smallest eigenvalues of $\Sigma_{X_1 X_1}$ and $\Sigma_{N_1 N_1}$.

In practical cases, non linearities and sampling error will add noise to the diagonal elements of $\hat H$. To test diagonal elements for equality with zero, we note that the average value of the non-zero coefficients of $\text{diag}(\hat H)$ is the trace of $\hat H$ divided by $k$. Since $X$ and $N$ are not correlated, $\Sigma_{X_1 X_1}^{-1}\Sigma_{N_1 N_1}$ should have a low condition number, which means $\frac{Var(X)}{Var(X)+Var(N)}$ should provide a good estimate of $\frac{1}{k}Tr(\hat H)$. This can be used to test whether a feature of $X$ has causal influence. An alternative is the Sonquist-Morgan criterion presented in the appendix.

@@

% I find a bit surprising to mention cauaslity literare in ML at the end. I would have expected this in the introduction to convince reviewer that know your stuff.

Finally, B2B is also related to, but different from causal discovery algorithms used outside the neuroimaging field. These methods are often nonlinear \citep{peters2017elements} and base their causal estimates on temporal delays \citep{granger1969investigating, janzing2013quantifying}. For example, B2B relates to the half-sibling regression \citep{scholkopf2016modeling}, in the sense that we observe a large number of variables corrupted by the same source of noise; we leave the extension of B2B to nonlinear time-series for future work.

### 2019-11-12 paper/synthetic.tex
new figs
correct alex comments

The data generating process for each experiment constructs $n=1000$ training examples
according to the model $Y = (\text{snr} \cdot XE + N)F$, where $\text{snr}$ is a
scalar that modulates the signal-to-noise ratio.
%
Here,
%\begin{itemize} \item $F \in \mathbb{R}^{d_x \times d_y}$ contains entries
%drawn from $\mathcal{N}(0, d_x^{-1})$, \item $X \in \mathbb{R}^{n \times d_x}$
%contains rows drawn from $\mathcal{N}(0, \Sigma_X)$, \item $N \in \mathbb{R}^{n
%\times d_x}$ contains rows drawn from $\mathcal{N}(0, \Sigma_N)$, \item $E \in
%\mathbb{R}^{d_x \times d_x}$ is a binary diagonal matrix containing $n_c$ ones,
%\item $\Sigma_X = AA^\top$, where $A \in \mathbb{R}^{d_x \times d_x}$ contains
%entries drawn from $\mathcal{N}(0, d_x^{-1})$, \item $\Sigma_N = BB^\top$,
%where $B \in \mathbb{R}^{d_x \times d_x}$ contains entries drawn from
%$\mathcal{N}(0, d_x^{-1})$, \end{itemize}
    $F \in \mathbb{R}^{d_x \times d_y}$ contains entries drawn from
$\mathcal{N}(0, d_x^{-1})$, $X \in \mathbb{R}^{n \times d_x}$ contains rows
drawn from $\mathcal{N}(0, \Sigma_X)$, $N \in \mathbb{R}^{n \times d_x}$
contains rows drawn from $\mathcal{N}(0, \Sigma_N)$, $E \in \mathbb{R}^{d_x
\times d_x}$ is a binary diagonal matrix containing $n_c$ ones, $\Sigma_X =
AA^\top$ where $A \in \mathbb{R}^{d_x \times d_x}$ contains entries drawn from
$\mathcal{N}(0, d_x^{-1})$, $\Sigma_N = BB^\top$ where $B \in \mathbb{R}^{d_x
\times d_x}$ contains entries drawn from $\mathcal{N}(0, d_x^{-1})$, and the
factor $\text{snr} \in (0, \infty)$.

To simulate a wide range of experimental conditions, we sample 10 values in log-space for $d_x, d_y \in \left[ 10, 100 \right]$, $n_c \in \left[ 3, 63 \right]$,
$\text{snr} \in \left[ 0.001, 10 \right]$. We discard the cases where $n_c > d_x$, limit $d_x, d_y$ to 100 to keep the running time under 2 hours for each condition, and average over 5 random seeds.
%
% Each condition is simulated under $5$ different random seeds.

@@

% To be checked
PLS finds $G\in\mathbb{R}^{d_z, d_y}$ and $H\in\mathbb{R}^{d_z, d_x}$
% such that
s.t.
$X$ and $Y$ are maximally covarying in a latent $Z$ space:
% \begin{equation} maxcov(XH^T, YG^T) \end{equation}
\begin{equation} G,H = \argmax_{G,H} cov(XH^T, YG^T) \end{equation}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{figures/AUC_conditions}
  \includegraphics[width=\linewidth]{figures/R_conditions}
  \vspace{-4ex}
  \caption{Synthetic experiments. Average AUC (top) and Feature Importance $\Delta R$ (bottom) when varying experimental conditions individually. Higher is better. B2B compares favorably in all cases. \label{fig:percondition}}
\end{figure}

>>>>

The data generating process for each experiment constructs $n=1000$ training examples
according to the model $Y = (\text{h} XE + N)F$, where $\text{h}$ is a
scalar that modulates the signal-to-noise ratio.
%
Here,
%\begin{itemize} \item $F \in \mathbb{R}^{d_x \times d_y}$ contains entries
%drawn from $\mathcal{N}(0, \sigma^2)$, \item $X \in \mathbb{R}^{n \times d_x}$
%contains rows drawn from $\mathcal{N}(0, \Sigma_X)$, \item $N \in \mathbb{R}^{n
%\times d_x}$ contains rows drawn from $\mathcal{N}(0, \Sigma_N)$, \item $E \in
%\mathbb{R}^{d_x \times d_x}$ is a binary diagonal matrix containing $n_c$ ones,
%\item $\Sigma_X = AA^\top$, where $A \in \mathbb{R}^{d_x \times d_x}$ contains
%entries drawn from $\mathcal{N}(0, \sigma^2)$, \item $\Sigma_N = BB^\top$,
%where $B \in \mathbb{R}^{d_x \times d_x}$ contains entries drawn from
%$\mathcal{N}(0, \sigma^2)$, \end{itemize}
    $F \in \mathbb{R}^{d_x \times d_y}$ contains entries drawn from
$\mathcal{N}(0, \sigma^2)$ where $\sigma^2$ is inversely proportional to $d_x$,
$X \in \mathbb{R}^{n \times d_x}$ contains rows
drawn from $\mathcal{N}(0, \Sigma_X)$, $N \in \mathbb{R}^{n \times d_x}$
contains rows drawn from $\mathcal{N}(0, \Sigma_N)$, $E \in \mathbb{R}^{d_x
\times d_x}$ is a binary diagonal matrix containing $n_c$ ones, $\Sigma_X =
AA^\top$ where $A \in \mathbb{R}^{d_x \times d_x}$ contains entries drawn from
$\mathcal{N}(0, \sigma^2)$, $\Sigma_N = BB^\top$ where $B \in \mathbb{R}^{d_x
\times d_x}$ contains entries drawn from $\mathcal{N}(0, \sigma^2)$, and the
factor $\text{h} \in \mathcal{R}_+$.

To simulate a wide range of experimental conditions, we sample 10 values in log-space for $d_x, d_y \in \left[ 10, 100 \right]$, $n_c \in \left[ 3, 63 \right]$,
$\text{h} \in \left[ 0.001, 10 \right]$. We discard the cases where $n_c > d_x$, limit $d_x, d_y$ to 100 to keep the running time under 2 hours for each condition, and average over 5 random seeds.
%
% Each condition is simulated under $5$ different random seeds.

@@

% To be checked
PLS finds $G\in\mathbb{R}^{d_z, d_y}$ and $H\in\mathbb{R}^{d_z, d_x}$
% such that
s.t.
$X$ and $Y$ are maximally covarying in a latent $Z$ space:
% \begin{equation} maxcov(XH^T, YG^T) \end{equation}
\begin{equation} G,H = \argmax_{G,H} \text{cov}(XH^T, YG^T) \end{equation}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{figures/auc_condition.pdf}
  \includegraphics[width=\linewidth]{figures/r_in_condition.pdf}
  \vspace{-4ex}
  \caption{Synthetic experiments. Average AUC (top) and Feature Importance $\Delta R$ (bottom) when varying experimental conditions individually. Higher is better. B2B compares favorably in all cases. \label{fig:percondition}}
\end{figure}
